---
title: "SDUSD Teacher Retention & CBS Strategy Report"
subtitle: "Retention Rate, Shifting Rate & Community-Based School (CBS) Strategy Trends"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    theme: flatly
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

<!--
PROJ: Teacher Retention
FILE: 05-retention-cbs-report.qmd
AUTH: Jeffrey Yo
INIT: 7/13/26

Goal: Pull together everything produced by 03-plot-retention-cbs-trends.R
(retention_cbs_plots.RData) and 04-correlation-analyses.R
(correlation_analysis_plots.RData + 5 CSVs) into one shareable HTML
report, organized as Overall / By Cohort / By School / Correlation
Analysis. embed-resources renders a single self-contained .html file -
safe to email or share without an accompanying assets folder.

Requires the DT package (install.packages("DT")) in addition to the
packages used by scripts 03/04.

Render with: quarto render 05-retention-cbs-report.qmd
-->

```{r setup}
library(tidyverse)
library(janitor)
library(scales)
library(DT)

data_dir<-file.path("C:/Users/jyo/Box/TR Data/SDUSD/original","data")
output_dir<-file.path("C:/Users/jyo/Box/TR Data/SDUSD/original","output")

#plot objects built by 03-plot-retention-cbs-trends.R and
#04-correlation-analyses.R
#load(file.path(data_dir, "retention_cbs_plots.RData"))

source(file.path(".","03-plot-retention-cbs-trends.R"))
load(file.path(data_dir, "correlation_analysis_plots.RData"))

#the retention/shifting plots store a sec_axis() formula
#(~ . * cbs_max) that references cbs_max from the ORIGINAL script's
#environment - that environment isn't saved along with the plot, so the
#lookup fails when the plot is printed in this fresh session unless we
#redefine cbs_max here too (must match the value used in
#03-plot-retention-cbs-trends.R)
cbs_max<-3

#numeric result tables written by 04-correlation-analyses.R
corr_table<-read_csv(file.path(output_dir, "correlation_table.csv"), show_col_types = FALSE)
cbs_stage_anova_results<-read_csv(file.path(output_dir, "cbs_stage_anova_results.csv"), show_col_types = FALSE)
cbs_stage_group_summary<-read_csv(file.path(output_dir, "cbs_stage_group_summary.csv"), show_col_types = FALSE)
cohort_anova_results<-read_csv(file.path(output_dir, "cohort_anova_results.csv"), show_col_types = FALSE)
cohort_group_summary<-read_csv(file.path(output_dir, "cohort_group_summary.csv"), show_col_types = FALSE)
lagged_results<-read_csv(file.path(output_dir, "lagged_correlation_results.csv"), show_col_types = FALSE)

#renders a data frame as a sortable/filterable table
dt_table<-function(df) {
  datatable(df, rownames = FALSE, filter = "top",
           options = list(pageLength = 10, scrollX = TRUE))
}
```

```{r lookups}
#the .RData plot lists are keyed by janitor::make_clean_names() slugs
#(e.g. "baker_elementary"), not the original display names - re-derive
#the slug -> display-name mapping the same way 03-plot-retention-cbs-
#trends.R does (same Marshall Elementary A/B fix), so tab and dropdown
#labels show the real school/cohort names

raw<-read_csv(file.path(data_dir, "school_retention_apr_long_2020_2024.csv"), show_col_types = FALSE) %>%
  clean_names()

raw<-raw %>%
  arrange(school_name, academic_year) %>%
  group_by(school_name, academic_year) %>%
  mutate(row_in_group = row_number()) %>%
  ungroup() %>%
  mutate(school_name = if_else(school_name == "Marshall Elementary",
                               paste("Marshall Elementary", if_else(row_in_group == 1, "A", "B")),
                               school_name)) %>%
  select(-row_in_group)

school_lookup<-tibble(school = sort(unique(raw$school_name))) %>%
  mutate(slug = make_clean_names(school))

ccspp_lookup<-tibble(cohort = sort(unique(raw$ccspp_cohort[!is.na(raw$ccspp_cohort)]))) %>%
  mutate(slug = make_clean_names(cohort))

sdusd_lookup<-tibble(cohort = sort(unique(raw$sdusd_cohort[!is.na(raw$sdusd_cohort)]))) %>%
  mutate(slug = make_clean_names(cohort))
```

This report summarizes teacher retention rate, teacher shifting rate, and Community-Based School (CBS) strategy scores for `r nrow(school_lookup)` schools across the 2020-21 through 2024-25 academic years. Use the **Overall**, **By Cohort**, and **By School** sections below to explore the trend plots, and the **Correlation Analysis** section for how retention/shifting relate to CBS strategies and other APR variables.

## Overall

::: {.panel-tabset}

### Retention Rate

```{r}
overall_plots$retention_rate
```

### Shifting Rate

```{r}
overall_plots$shifting_rate
```

### Combined

```{r}
overall_plots$combined
```

:::

## By Cohort

### CCSPP Cohort

```{r}
#| output: asis
for (i in seq_len(nrow(ccspp_lookup))) {

  coh<-ccspp_lookup$cohort[i]
  slug<-ccspp_lookup$slug[i]

  cat("#### ", coh, "\n\n")
  cat("::: {.panel-tabset}\n\n")

  cat("##### Retention Rate\n\n")
  print(ccspp_cohort_plots$retention_rate[[slug]])

  cat("\n\n##### Shifting Rate\n\n")
  print(ccspp_cohort_plots$shifting_rate[[slug]])

  cat("\n\n##### Combined\n\n")
  print(ccspp_cohort_plots$combined[[slug]])

  cat("\n\n:::\n\n")
}
```

### SDUSD Cohort

```{r}
#| output: asis
for (i in seq_len(nrow(sdusd_lookup))) {

  coh<-sdusd_lookup$cohort[i]
  slug<-sdusd_lookup$slug[i]

  cat("#### ", coh, "\n\n")
  cat("::: {.panel-tabset}\n\n")

  cat("##### Retention Rate\n\n")
  print(sdusd_cohort_plots$retention_rate[[slug]])

  cat("\n\n##### Shifting Rate\n\n")
  print(sdusd_cohort_plots$shifting_rate[[slug]])

  cat("\n\n##### Combined\n\n")
  print(sdusd_cohort_plots$combined[[slug]])

  cat("\n\n:::\n\n")
}
```

## By School

Pick a school from the dropdown to see its retention/shifting/combined trend plots.

```{=html}
<style>
.school-panel { display: none; }
.school-panel.active { display: block; }
.school-selector-wrapper { margin: 1em 0 1.5em 0; }
#school-select { font-size: 1rem; padding: 0.35em; min-width: 260px; }
</style>
<script>
function showSchoolPanel(slug) {
  var panels = document.getElementsByClassName('school-panel');
  for (var i = 0; i < panels.length; i++) {
    panels[i].classList.remove('active');
  }
  var target = document.getElementById('school-panel-' + slug);
  if (target) { target.classList.add('active'); }
}
</script>
```

```{r}
#| output: asis
cat('<div class="school-selector-wrapper">\n')
cat('<label for="school-select"><strong>Select a school: </strong></label>\n')
cat('<select id="school-select" onchange="showSchoolPanel(this.value)">\n')
for (i in seq_len(nrow(school_lookup))) {
  cat(sprintf('<option value="%s">%s</option>\n', school_lookup$slug[i], school_lookup$school[i]))
}
cat('</select>\n')
cat('</div>\n\n')
```

```{r}
#| output: asis
for (i in seq_len(nrow(school_lookup))) {

  sch<-school_lookup$school[i]
  slug<-school_lookup$slug[i]
  active_class<-if (i == 1) " active" else ""

  cat(sprintf('<div id="school-panel-%s" class="school-panel%s">\n\n', slug, active_class))

  cat("#### ", sch, "\n\n")
  cat("::: {.panel-tabset}\n\n")

  cat("##### Retention Rate\n\n")
  print(school_plots$retention_rate[[slug]])

  cat("\n\n##### Shifting Rate\n\n")
  print(school_plots$shifting_rate[[slug]])

  cat("\n\n##### Combined\n\n")
  print(school_plots$combined[[slug]])

  cat("\n\n:::\n\n")
  cat('</div>\n\n')
}
```

## Correlation Analysis

How retention rate relates to the 5 CBS strategy scores and other numeric APR/participation variables. See `04-correlation-analyses.R` for the underlying methodology. Shifting rate was dropped from this analysis (near-zero variation across schools made it uninformative to correlate against).

### Correlation Matrix

```{r}
corr_heatmap
```

Significant cells (p < 0.05) are starred.

**Full correlation table** (Pearson and Spearman, every predictor x outcome):

```{r}
dt_table(corr_table)
```

### Scatterplots

One panel per CBS strategy score (plus teacher leadership development support), with a linear trend line. Each panel is labeled with its Pearson r, starred when p < 0.05.

```{r}
#| fig-height: 11
scatter_plots$retention_rate
```

### CBS Maturity Stage Comparison (ANOVA)

Retention rate compared across the 3 CBS maturity stages (1 - Visioning, 2 - Engaging, 3 - Transforming) for each CBS strategy, via one-way ANOVA.

```{r}
cbs_stage_boxplots$retention_rate
```

**ANOVA results** (retention rate ~ stage, per CBS strategy):

```{r}
dt_table(cbs_stage_anova_results)
```

**Group means by stage:**

```{r}
dt_table(cbs_stage_group_summary)
```

### Cohort Comparisons (ANOVA)

::: {.panel-tabset}

#### CCSPP Cohort

```{r}
cohort_plots$ccspp_cohort$retention_rate
```

#### SDUSD Cohort

```{r}
cohort_plots$sdusd_cohort$retention_rate
```

:::

**ANOVA results:**

```{r}
dt_table(cohort_anova_results)
```

**Group means by cohort:**

```{r}
dt_table(cohort_group_summary)
```

### Lagged Correlation

Spearman correlation of this year's CBS score against the *same*-year retention rate ("lag 0") vs. *next*-year retention rate ("lag 1"), within school.

```{r}
lagged_plots$retention_rate
```

**Lagged correlation results:**

```{r}
dt_table(lagged_results)
```
